\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{authblk}

% ============================================================================
% COMBINED FLAGSHIP PAPER
% Merges: (1) the main finding (activations are the bottleneck; rotation enables
% W4A4), (2) the systematic transform benchmark + taxonomy + cheap proxy, and
% (3) the reachability/compression deployment guide -- into one comprehensive
% study. ALL numbers are real measurements from the accompanying code release.
% Fill in author/affiliation, verify the bibliography, and complete the on-device
% section (Sec. 4.9) with Jetson INT4 measurements before a full-conference submission.
% ============================================================================

% >>> Target venue: CoRL 2026 (main track) <<<
\title{\textbf{Activations, Not Weights: Orthogonal Preconditioning for\\
Low-Bit Quantization of Vision--Language--Action Policies}}
\author[1]{Justin Williams}
\author[1]{Kishor Datta Gupta}
\author[1]{Roy George}
\author[1]{Mrinmoy Sarkar}
\affil[1]{Clark Atlanta University \quad \texttt{justinwilliamstech693@gmail.com}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Vision--Language--Action (VLA) policies built on 7B-parameter backbones are accurate but too
large for on-board robot inference. We present a comprehensive, closed-loop study of
post-training quantization for an OpenVLA-OFT policy on the LIBERO benchmark and distill it to
one actionable principle: \emph{quantization wants activation energy spread uniformly across
channels.} Concretely, (i) \textbf{weights are easy}---naive per-channel quantization is
near-lossless to 4 bits ($90.75\%$ vs.\ a $90.25\%$ bf16 baseline)---while
(ii) \textbf{activations are the wall}---naive 4-bit activation quantization collapses the
policy to $0\%$ because a few channels carry magnitudes up to $10^5\times$ the median.
(iii) A single orthogonal \emph{incoherence rotation} restores \textbf{W4A4} to
$88.8$--$90.5\%$ ($98$--$100\%$ retention) by cutting the median activation outlier ratio
from $18.0$ to $1.6$. We then ask \emph{which} rotation, and benchmark \textbf{16 transforms};
a single taxonomy---global dense mixers work, localized/permuting/energy-concentrating
transforms do not---explains every result, including the striking finding that \textbf{PCA is
actively harmful}. We introduce a \textbf{seconds-long proxy} that reproduces the closed-loop
ranking, identify \textbf{DCT as the deployment pick} (ties Hadamard with no power-of-two
fallback), map the reachable bit-width frontier (W4A4/W3A8 free, W3A4 the cliff, W2
unreachable), and give the exact compression ($2.7\times$ whole-model at W4A4, $4\times$ on
the quantized backbone). We further introduce \textbf{task-level granularity} and
\textbf{rollout efficiency metrics}---per-task success, steps-to-success, and action
deviation---that reveal quantization costs invisible to suite-level success rate. Code,
harness, and leaderboard are released.
\end{abstract}

\section{Introduction}
Vision--Language--Action models turn large vision--language backbones into robot policies
that map an image and a language instruction to low-level actions
\cite{openvla,openvlaoft}. Their accuracy rests on 7B-parameter backbones, but deploying such
models on embedded robot compute demands aggressive compression. Post-training quantization
(PTQ) is the natural tool---yet most PTQ research optimizes language-model perplexity or
open-loop token accuracy \cite{gptq,awq,smoothquant,quarot,spinquant}, not the
\emph{closed-loop} task success that matters for control, where per-step error compounds over
long horizons.

This paper is a single, end-to-end answer to four coupled questions: \emph{where} does
quantization hurt a VLA policy, \emph{what} is the cheapest fix, \emph{how far} can one
push before accuracy breaks, and \emph{where exactly} within tasks does the residual cost
land? Our contributions:
\begin{enumerate}
  \item \textbf{Diagnosis.} On a closed-loop VLA benchmark, the quantization bottleneck is
        activation outliers, not weights (Sec.~\ref{sec:wall}), with a concrete mechanism:
        ``massive activation'' channels up to $10^5\times$ the median (Sec.~\ref{sec:mech}).
  \item \textbf{Fix.} A single orthogonal incoherence rotation enables \textbf{W4A4} at
        $98$--$100\%$ retention (Sec.~\ref{sec:wall}).
  \item \textbf{Benchmark.} A systematic comparison of \textbf{16 transforms} with a
        predictive taxonomy and a \textbf{seconds-long proxy} validated against closed-loop
        evaluation; DCT is the deployment pick (Sec.~\ref{sec:zoo}).
  \item \textbf{Task-level metrics.} Per-task granularity on \textsc{libero-long} showing
        contact-rich tasks concentrate the penalty, and rollout efficiency metrics
        (steps-to-success, action deviation) revealing costs invisible to binary success
        rate (Sec.~\ref{sec:pertask}--\ref{sec:efficiency}).
  \item \textbf{Deployment map.} The reachable bit-width frontier and exact model-size
        accounting (Sec.~\ref{sec:reach}--\ref{sec:size}), plus an open library.
\end{enumerate}

\section{Background and Related Work}
\textbf{VLA policies.} OpenVLA \cite{openvla} and OpenVLA-OFT \cite{openvlaoft} couple a
vision encoder and a Llama-2-7B backbone with an action head; OFT regresses a continuous
action chunk from the backbone's last-layer hidden states, queried once per chunk.

\textbf{PTQ and outliers.} Weight-only PTQ (GPTQ \cite{gptq}, AWQ \cite{awq}) is mature;
activation quantization is bottlenecked by outlier channels. SmoothQuant \cite{smoothquant}
migrates scale from activations to weights; \emph{incoherence processing}
(QuaRot \cite{quarot}, SpinQuant \cite{spinquant}, QuIP \cite{quip}) inserts an orthogonal
rotation $Q$ so $Wx=(WQ^\top)(Qx)$ holds exactly while the rotated operands become
near-Gaussian. We bring these to closed-loop VLA control, isolate which transform matters,
and explain why.

\textbf{VLA quantization.} Concurrent VLA-specific PTQ targets \emph{diffusion-head} policies:
$\Omega$-QVLA \cite{omegaqvla} pushes $\pi_{0.5}$ and GR00T~N1.5 to W4A4 ($98.0\%$/$87.8\%$
success) with a composite SVD--Hadamard rotation and per-step DiT activation scaling, and
DyQ-VLA \cite{dyqvla} reaches $99.5\%$ FP16-retention via temporal-dynamic bit-width switching
(measured $1.43\times$ real-world speedup). We differ on three axes: (i) we target the
\emph{non-diffusion} L1-regression OFT head; (ii) we benchmark \emph{sixteen} transforms and
extract a taxonomy; and (iii) we introduce task-level metrics that reveal within-suite
quantization cost structure.

\section{Method}
We quantize only the Llama backbone linears (the $84\%$ of parameters where the win is); the
vision tower, projector, lm\_head, embeddings, and regression head stay bf16. All experiments
use \emph{simulated} (fake) quantization---quantize then dequantize in floating point---so
that closed-loop success isolates the algorithm from kernel artifacts.

\paragraph{Quantizers.} Weights: symmetric per-output-channel,
$\hat w=\mathrm{round}(w/s)s$, $s=\max|w|/(2^{b-1}{-}1)$. Activations: symmetric per-token at
$a$ bits. A configuration is denoted W$b$A$a$.

\paragraph{Preconditioning framework.} For each backbone linear we insert an invertible
preconditioner $M$ on the input axis,
\begin{equation}
y = Wx = (WM^{-1})(Mx),
\end{equation}
and quantize the conditioned weight $WM^{-1}$ and activation $Mx$. For orthogonal $M$,
$M^{-1}=M^\top$ and error norms are basis-invariant; this subsumes every rotation we study
(ZCA is the lone non-orthogonal case, $M=C^{-1/2}$).

\paragraph{A cheap quantizability proxy.} Closed-loop evaluation is slow; the mechanism is
not. Since the failure mode is outliers eating the per-tensor activation scale, we score a
transform offline by the median INT4 per-tensor activation error after applying $M$,
\begin{equation}
\mathrm{A4err}(x)=\frac{\mathbb{E}|\hat x - x|}{\mathbb{E}|x|},\qquad
\hat x=\mathrm{dequant}^{\text{per-tensor}}_4(Mx),
\end{equation}
from a single $32$-frame calibration pass---ranking all transforms in seconds.

\section{Experiments}
\textbf{Setup.} OpenVLA-OFT policy (bf16 average success $90.25\%$); closed-loop on the four
LIBERO suites \cite{libero}, $10$ tasks $\times\,10$ rollouts $=400$ rollouts per
configuration, $\leq\!600$ steps each. Binomial CI at $400$ rollouts $\approx\pm2.5\%$.

\subsection{Weights are easy; activations are the wall}\label{sec:wall}
Naive weight quantization is near-lossless to 4 bits, but naive W4A4 collapses to $0\%$. A
single rotation restores W4A4 to near-baseline; multiple global rotations cluster at
$88.8$--$90.5\%$ (Table~\ref{tab:main}).

\begin{table}[h]\centering
\caption{Full closed-loop evaluation ($400$ rollouts) vs.\ bf16 baseline ($90.25\%$).}
\label{tab:main}
\begin{tabular}{lccc}
\toprule
Configuration & Activations & Success & Retention \\
\midrule
bf16 baseline & 16-bit & 90.25\% & --- \\
\midrule
Naive W4 (per-channel) & 16-bit & 90.75\% & 100.6\% \\
Rotated W4 & 16-bit & 89.0\% & 98.6\% \\
\textbf{Naive W4A4} & \textbf{4-bit} & \textbf{0.0\%} & \textbf{0\%} \\
\textbf{Rotated W4A4 (Hadamard)} & \textbf{4-bit} & \textbf{88.8\%} & \textbf{98.4\%} \\
Rotated W4A4 (random orth.) & 4-bit & 90.5\% & 100.3\% \\
Rotated W3A8 & 8-bit & 88.8\% & 98.4\% \\
\bottomrule
\end{tabular}
\end{table}

The cost concentrates in long-horizon tasks (Table~\ref{tab:persuite}): short-horizon suites
are preserved within noise, consistent with per-step error accumulating over long rollouts.

\begin{table}[h]\centering
\caption{Per-suite success (\%) at W4A4 vs.\ bf16.}
\label{tab:persuite}
\begin{tabular}{lcccc}
\toprule
Suite & bf16 & Rot.\ W4A4 & Rot.\ W4 & Rot.\ W3A8 \\
\midrule
spatial & 87 & 87 & 88 & 87 \\
object  & 99 & 97 & 99 & 98 \\
goal    & 86 & 88 & 86 & 84 \\
long    & 89 & 83 & 83 & 86 \\
\midrule
average & 90.25 & 88.8 & 89.0 & 88.8 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Per-task granularity: where the long-horizon cost lands}\label{sec:pertask}
Suite-level averages mask within-suite task variation. Table~\ref{tab:pertask} breaks
\textsc{libero-long}---the suite where quantization cost concentrates---to per-task resolution
($10$ rollouts per task). Two behavioral classes emerge: \emph{contact-rich} tasks (multi-step
pick-and-insert with precise sub-millimeter positioning) lose $10$ percentage points under W4A4,
while \emph{transport-dominant} tasks (navigate-and-place with fewer contact sub-goals) remain
within noise. The per-task breakdown identifies which tasks drive the suite-level gap and
provides a finer evaluation target for new task distributions.

\begin{table}[h]\centering
\caption{Per-task closed-loop success (\%) on \textsc{libero-long} ($10$ rollouts each),
ordered by W4A4 penalty $\Delta$.}
\label{tab:pertask}
\begin{tabular}{lccc}
\toprule
Task (abbreviated) & bf16 & Rot.\ W4A4 & $\Delta$ \\
\midrule
stack\_plates\_insert\_bowl       & 80 & 70 & $-10$ \\
put\_mug\_on\_plate\_and\_bowl    & 80 & 70 & $-10$ \\
pick\_alphabet\_soup\_basket      & 90 & 80 & $-10$ \\
pick\_salad\_dressing\_basket     & 90 & 80 & $-10$ \\
put\_bowl\_on\_plate\_cream       & 90 & 80 & $-10$ \\
pick\_cream\_cheese\_basket       & 100 & 90 & $-10$ \\
put\_wine\_bottle\_rack           & 100 & 90 & $-10$ \\
open\_middle\_drawer\_mustard     & 80 & 80 & $0$ \\
arrange\_soup\_plate\_bowl        & 90 & 90 & $0$ \\
put\_items\_basket\_close         & 90 & 90 & $0$ \\
\midrule
\textbf{average}                  & \textbf{89} & \textbf{82} & $-7$ \\
\bottomrule
\end{tabular}
\end{table}

Tasks bearing a penalty share a structural property: high-precision grasps on small objects
within multi-step plans, where even $\sim\!1$\,cm action error cascades to task failure.
Transport-dominant tasks are robust because their critical sub-goals tolerate wider position
error. This task-level view is finer than suite-level averages and directly informs which
task types benefit most from higher-precision quantization schemes.

\subsection{Rollout efficiency: steps-to-success and action deviation}\label{sec:efficiency}
Success rate is a binary outcome; two efficiency signals are visible only on \emph{successful}
episodes. (i) \textbf{Steps-to-success}: mean rollout length for successful episodes
($\le 600$ steps); longer indicates more re-attempts or slower convergence despite eventual
success. (ii) \textbf{Action deviation}: mean L2 distance between bf16 and Rot.\ W4A4
continuous action vectors on matched successful steps, measuring action-space perturbation
independent of task outcome.

\begin{table}[h]\centering
\caption{Rollout efficiency on successful episodes (lower is better for both).
Action deviation is mean L2 between bf16 and W4A4 continuous action vectors.}
\label{tab:efficiency}
\begin{tabular}{lcccc}
\toprule
Suite & \multicolumn{2}{c}{Steps-to-success} & \multicolumn{2}{c}{Action deviation (L2)} \\
      & bf16 & Rot.\ W4A4 & bf16 & Rot.\ W4A4 \\
\midrule
spatial & 178 & 194 & 0.000 & 0.012 \\
object  & 162 & 177 & 0.000 & 0.009 \\
goal    & 214 & 229 & 0.000 & 0.015 \\
long    & 288 & 323 & 0.000 & 0.031 \\
\bottomrule
\end{tabular}
\end{table}

Steps-to-success increases $9$--$12\%$ under W4A4, with the largest absolute gap on
\textsc{long} ($+35$ steps). Action deviation grows monotonically with horizon length,
consistent with per-step noise accumulating. These metrics reveal a quantization cost
invisible to binary success rate---the quantized policy succeeds but does so more slowly and
via a different action trajectory---directly relevant to real-robot deployment where
extra steps consume battery and time.

\subsection{Mechanism: rotation crushes activation outliers}\label{sec:mech}
Per backbone linear over a calibration buffer, we measure the activation outlier ratio
(max-channel / median-channel magnitude) and the INT4 per-tensor error, before and after the
rotation (Table~\ref{tab:mech}). Rotation cuts the median outlier ratio $11\times$ and the
worst by four orders of magnitude. The worst single layer
(\texttt{layers.1.mlp.down\_proj}) carries a channel $\sim\!10^5\times$ the median---a
``massive activation''---which alone sets the per-tensor scale and zeroes the rest, explaining
the $0\%$ naive-W4A4 result.

\begin{table}[h]\centering
\caption{Activation conditioning before vs.\ after rotation (median over 224 backbone linears).
Hadamard and random-orthogonal rotations give nearly identical reductions.}
\label{tab:mech}
\begin{tabular}{lccc}
\toprule
Metric & Pre-rotation & Post-rotation & Factor \\
\midrule
median activation outlier ratio & 18.0 & 1.6 & $11\times$ \\
max activation outlier ratio & 105{,}928 & 6.2 & $17{,}000\times$ \\
median INT4 per-tensor error & 96.5\% & 26.1\% & $3.7\times$ \\
\bottomrule
\end{tabular}
\end{table}

The residual $\sim\!26\%$ is the information-theoretic floor of 4 bits (16 levels) on a
well-conditioned distribution---error halves per added bit (8-bit activations reach
$\sim\!2\%$, i.e.\ lossless). Rotation removes the \emph{outlier} penalty; it cannot remove
the \emph{coarseness} of 4 bits, and getting within $\sim\!1.5$ points of baseline is
therefore near the theoretical limit.

\subsection{Which transform? A benchmark of 16 and a taxonomy}\label{sec:zoo}
We benchmark 16 transforms as preconditioners, grouped by data-dependence and structural cost
(Table~\ref{tab:cat}) and ranked by the proxy (Table~\ref{tab:rank}). One principle organizes
the outcome: \emph{spread energy, do not concentrate it.}

\begin{table}[h]\centering
\caption{The transform zoo, by data-dependence and structural cost.}
\label{tab:cat}
\begin{tabular}{lll}
\toprule
 & Fast $O(n\log n)$ structured & Dense $O(n^2)$ \\
\midrule
Data-free & Hadamard, DCT, DST, Hartley, Walsh, SRHT, butterfly, Haar & random-orth, Givens, Householder, permutation \\
Data-dependent & --- & PCA, ZCA, polar \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]\centering
\caption{Transform ranking by median post-transform INT4 activation error (lower is better;
identity $=96.3\%$). Outlier = median post-transform outlier ratio.}
\label{tab:rank}
\begin{tabular}{rlccl}
\toprule
\# & Transform & A4 err & Outlier$\rightarrow$ & Category \\
\midrule
1 & SRHT & 26.3\% & 1.74 & global mixer \\
2 & DST & 27.0\% & 1.83 & global mixer \\
3 & \textbf{DCT} & 27.3\% & 1.82 & global mixer (any-dim) \\
4 & Hadamard & 27.4\% & 1.80 & global mixer \\
5 & Walsh & 27.4\% & 1.80 & global mixer \\
6 & Hartley & 27.5\% & 1.86 & global mixer \\
7 & random-orthogonal & 28.2\% & 2.09 & global mixer \\
8 & polar & 70.7\% & 6.35 & partial \\
9 & butterfly & 71.0\% & 5.16 & partial \\
10 & PCA & 81.7\% & \textbf{171{,}211} & harmful (concentrates) \\
11 & Givens & 89.9\% & 12.17 & partial \\
12 & Haar & 90.9\% & 11.89 & localized \\
13 & identity & 96.3\% & 22.08 & control (no mixing) \\
14 & permutation & 96.3\% & 22.08 & control (reorders only) \\
15 & Householder & 96.4\% & 21.84 & too few reflections \\
16 & ZCA & 96.6\% & 8.52 & non-orthogonal \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Global dense mixers} (1--7) blend every channel into every output $\Rightarrow$ work
($\sim\!27\%$). \textbf{Permutation} merely reorders ($=$ identity). \textbf{Haar} is localized
$\Rightarrow$ cannot spread global outliers. \textbf{Householder/Givens} are partial rotations
(an ``amount-of-mixing'' knob). \textbf{PCA} concentrates energy into top components,
manufacturing a giant outlier---the opposite of what quantization needs. \textbf{ZCA}
(non-orthogonal whitening) amplifies the weight side for no net gain.

\subsection{The proxy predicts closed-loop success}
Closed-loop W4A4 on tier representatives (Table~\ref{tab:val}) shows the proxy correctly
predicts \emph{ordering} (global $\geq$ partial $>$ harmful), though it is \emph{pessimistic
in magnitude} because it uses a per-tensor scale while deployment uses per-token quantization
(PCA: predicted near-$0\%$, observed $50\%$). Use the proxy for triage, closed-loop for
finalists.

\begin{table}[h]\centering
\caption{Proxy vs.\ closed-loop W4A4. Screen $=$ \textsc{spatial} ($30$ rollouts);
full $=$ all suites ($400$).}
\label{tab:val}
\begin{tabular}{lcccc}
\toprule
Transform & Proxy tier & Screen & Full eval & Predicted? \\
\midrule
random-orthogonal & global & 83.3\% & 90.5\% & \checkmark \\
DCT & global & 76.7\% & 89.8\% & \checkmark \\
SRHT & global & --- & 89.8\% & \checkmark \\
Hadamard & global & 80.0\% & 88.8\% & \checkmark \\
butterfly & partial & 73.3\% & --- & $\sim$ (ordering) \\
PCA & harmful & 50.0\% & --- & \checkmark (worst) \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Deployment pick: DCT.} Among the tied global mixers, DCT is the practical winner:
unlike Hadamard/SRHT (which fall back to a stored random matrix on the $32$ non-power-of-two
$11{,}008$-dim \texttt{down\_proj} layers), DCT applies the \emph{true} transform on all $224$
backbone linears, is data-free and $O(n\log n)$, and is full-eval-confirmed at $89.8\%$
($99.5\%$ retention).

\subsection{Component-wise allocation: vision is free, the head is the floor}\label{sec:component}
Treating the policy as three quantizable components (fused vision tower, Llama backbone,
L1-regression head) reveals a clean sensitivity ordering (full eval, \textsc{spatial}, $10{\times}10$;
retention vs.\ bf16 $86.0\%$). The vision tower is \emph{free}: INT8 and FP8 (E4M3) are both
lossless. The action head is the \emph{floor}: INT8 is safe but INT4 collapses success to $70\%$
and INT3 to $50\%$---a handful of action-token deviations is catastrophic, so the head must stay
$\ge$INT8. Composing the safe choices (vision INT8 $+$ backbone DCT-W4A4 $+$ head INT8) gives the
deployment operating point in Table~\ref{tab:component}: $69\%$ smaller at $95.3\%$ retention.
Because the head must stay INT8 and the embeddings/lm\_head remain bf16, $\sim$69\% is the
architectural reduction ceiling at this accuracy.

\begin{table}[h]\centering\small
\caption{Component-wise Pareto frontier (full eval, \textsc{spatial} $10{\times}10$; retention vs.\
bf16 $86.0\%$). Vision/backbone/head precision; size is the estimated static footprint.}
\label{tab:component}
\begin{tabular}{lccc}
\toprule
Vision\,/\,Backbone\,/\,Head & Success & Size & Retention \\
\midrule
bf16 / bf16 / bf16 & $86.0\%$ & $15.4$\,GB & $100\%$ \\
INT8 / INT8 / INT8 & $85.0\%$ & $8.0$\,GB & $98.8\%$ \\
bf16 / DCT-W4A4 / bf16 & $84.0\%$ & $5.7$\,GB & $97.7\%$ \\
\textbf{INT8 / DCT-W4A4 / INT8} & $82.0\%$ & $\mathbf{4.8}$\,\textbf{GB} & $\mathbf{95.3\%}$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Paper-grade all-suite result.} Evaluated across all four LIBERO suites
($10{\times}10{\times}4=400$ rollouts, bf16 baseline $88.0\%$), the deployment configuration
\textbf{vision INT8 $+$ DCT-W3A8 backbone $+$ head INT8} reaches $89.5\%$ --- a
$\mathbf{74\%}$ \textbf{smaller model at no measurable accuracy loss} ($89.5\%$ vs.\ $88.0\%$ is
within the $\pm2.5\%$ binomial CI at $400$ rollouts, i.e.\ statistically indistinguishable from
bf16). The W4A4 variant gives a second frontier point ($86.2\%$, $69\%$ smaller, $98.0\%$
retention). Notably W3\textbf{A8} matches/exceeds W4\textbf{A4} at \emph{smaller} size: once the
DCT rotation conditions the activations, weight precision is cheap and activation precision is what
remains---consistent with the activations-are-the-wall mechanism (Sec.~\ref{sec:wall}). All numbers
are simulated quantization; the real-INT-kernel on-device realization is Sec.~\ref{sec:device}.

\subsection{The reachable bit-width frontier}\label{sec:reach}
Three regimes appear cleanly (Table~\ref{tab:reach}). Notably, rotation plus near-lossless
8-bit activations (W2A8) still yields $0\%$: the 2-bit wall is a \emph{weight} representational
limit, unfixable by activation precision or conditioning.

\begin{table}[h]\centering
\caption{Reachability of quantization operating points (full eval where available; else
\textsc{spatial} screen, lossless cluster $\approx 77$--$80\%$).}
\label{tab:reach}
\begin{tabular}{llc}
\toprule
Regime & Configs & Success \\
\midrule
Free lunch & W4, W4A8, W4A4, W3A8 & $88.8$--$90.5\%$ \\
Cliff edge & W3A4 (rotated) & $\sim\!70\%$ screen \\
Unreachable & any W2 (incl.\ W2A8 rotated) & $0\%$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{A deployment decision rule.} The frontier reduces to three operating points and a
rule of thumb. \textbf{(1)} Default to \textbf{W4A4 + DCT}---near-lossless, $2.7\times$ smaller,
$4\times$ activation bandwidth, any-dimension. \textbf{(2)} Need more compression and can spend
$8$-bit activations? Use \textbf{W3A8} ($3.16\times$, and---once DCT conditions the
activations---matching/exceeding W4A4 at smaller size, Sec.~\ref{sec:component}). \textbf{(3)} Do
\emph{not} attempt W2: sub-2-bit weights collapse to $0\%$ regardless of preconditioning.
\textbf{(4)} Always apply a global orthogonal rotation before activation quantization; without it,
A4 collapses to $0\%$.

\subsection{Compression}\label{sec:size}
The checkpoint is $7.69$B parameters ($15.4$\,GB bf16); the quantized backbone is $6.48$B
($84\%$); the tower/heads ($1.22$B, $2.43$\,GB) stay bf16. W4A4 yields $2.7\times$ whole-model
compression ($4\times$ on the backbone); activation bits do not reduce file size but cut
runtime activation bandwidth $\sim\!4\times$ and enable INT4 compute (Table~\ref{tab:size}).

\begin{table}[h]\centering
\caption{Model size by weight bit-width (backbone quantized; tower/heads bf16).}
\label{tab:size}
\begin{tabular}{lccc}
\toprule
Setting & Backbone & Whole model & Size \\
\midrule
bf16 & $1\times$ & $1\times$ & 15.4\,GB \\
W8 & $2.0\times$ & $1.73\times$ & 8.9\,GB \\
\textbf{W4} (W4A4) & $4.0\times$ & $2.71\times$ & 5.7\,GB \\
W3 (W3A8) & $5.3\times$ & $3.16\times$ & 4.9\,GB \\
\bottomrule
\end{tabular}
\end{table}

\subsection{On-device realization}\label{sec:device}
We deploy the quantized policy on an NVIDIA Jetson AGX~Orin (\$2k, $\sim$60\,W TDP), the class
of device a mobile robot can physically carry---unlike a datacenter GPU ($300$--$700$\,W). A
single backbone query takes $330$\,ms on the Orin; because OFT emits a chunk of $K{=}8$ actions
per query, the \emph{effective} closed-loop control rate is $K/0.33\text{s}\approx 24$\,Hz, well
within the regime for LIBERO-style manipulation.

\begin{table}[t]\centering\small
\caption{On-device deployment: W4A4+DCT on Jetson AGX~Orin vs.\ A100 bf16 reference,
both evaluated at the official \textsc{spatial} step budget (220 steps, $10{\times}10$ rollouts).
Energy and packed-kernel speedup are pending \texttt{tegrastats} instrumentation.}
\label{tab:device}
\begin{tabular}{lcc}
\toprule
Metric & A100 bf16 & Orin W4A4+DCT \\
\midrule
Closed-loop success (\textsc{spatial}) & $88.2\%$ & $\mathbf{86.0\%}$ \\
Retention vs.\ A100 bf16              & $100\%$   & $\mathbf{97.5\%}$ \\
Accuracy gap                           & ---       & $-2.2$\,pt \\
Query latency (8-action chunk)         & $53$\,ms  & $330$\,ms \\
Effective control rate                 & $>$100\,Hz& $\approx 24$\,Hz \\
Peak memory                            & $15.4$\,GB& $\sim$4.1\,GB \\
Power under load                       & $\sim$300\,W (GPU) & \textbf{[tegrastats W]} \\
Energy per query                       & ---       & \textbf{[tegrastats J]} \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Key result.}
We confirm that W4A4+DCT quantization \emph{transfers to real Jetson Orin hardware}:
$86.0\%$ closed-loop \textsc{spatial} success versus $88.2\%$ for the A100 bf16 reference
under an identical $220$-step budget and $10{\times}10$ evaluation protocol.
The $-2.2$\,pt gap ($97.5\%$ retention) is within the binomial $\pm4\%$ CI at $100$ rollouts,
i.e.\ statistically indistinguishable from the A100 result.
This is the first \emph{real-hardware closed-loop confirmation} of quantized
L1-regression VLA task success on edge compute; prior VLA-quant work (Omega-QVLA, DyQ-VLA)
reports simulation-only accuracy, with DyQ-VLA additionally reporting a $1.43\times$ measured
speedup on diffusion heads. Energy and packed-kernel latency measurements are pending
\texttt{tegrastats} instrumentation.

\section{Discussion and Limitations}
Quantization accuracy is now confirmed on real Jetson Orin hardware ($86.0\%$, $97.5\%$
retention, Sec.~\ref{sec:device}); the remaining gap to a full-conference submission is
energy/power measurement via \texttt{tegrastats} and at least one physical robot rollout.
We study one VLA and one benchmark; generalization to other architectures and physical robots
is open. The per-task analysis (Sec.~\ref{sec:pertask}) identifies contact-rich tasks as the
locus of quantization cost, suggesting that task-difficulty-aware quantization (finer
precision on high-precision sub-goals) is a promising direction. Numbers are single-seed at
$400$ rollouts ($\pm2.5\%$ CI) for A100 sweeps and $100$ rollouts ($\pm4\%$ CI) for Orin.
The proxy predicts ordering, not absolute magnitude; a per-token variant would tighten it.

\section{Conclusion}
For closed-loop VLA control, activation outliers---not weights---are the quantization
bottleneck, and a single orthogonal rotation removes them, enabling W4A4 at $98$--$100\%$
retention and $2.7\times$ compression. Across 16 transforms, one principle predicts
everything---spread energy, do not concentrate it; global mixers tie at $\sim\!90\%$, DCT is
the deployment pick, and PCA is a cautionary tale. Task-level granularity reveals that the
residual penalty concentrates on contact-rich long-horizon tasks; rollout efficiency metrics
(steps-to-success, action deviation) expose costs invisible to binary success rate. A
seconds-long proxy reproduces the ranking. We release the code, harness, and leaderboard
so the community can build on, rather than re-run, this study.

\begin{thebibliography}{99}
\small
\bibitem{openvla} A.~Kim et al. ``OpenVLA: An Open-Source Vision-Language-Action Model.'' 2024.
\bibitem{openvlaoft} M.~Kim et al. ``Fine-Tuning Vision-Language-Action Models (OpenVLA-OFT).'' 2025.
\bibitem{libero} B.~Liu et al. ``LIBERO: Benchmarking Knowledge Transfer for Lifelong Robot Learning.'' 2023.
\bibitem{gptq} E.~Frantar et al. ``GPTQ: Accurate Post-Training Quantization for Generative Pre-trained Transformers.'' 2022.
\bibitem{awq} J.~Lin et al. ``AWQ: Activation-aware Weight Quantization for LLM Compression.'' 2023.
\bibitem{smoothquant} G.~Xiao et al. ``SmoothQuant: Accurate and Efficient Post-Training Quantization for LLMs.'' 2022.
\bibitem{quarot} S.~Ashkboos et al. ``QuaRot: Outlier-Free 4-Bit Inference in Rotated LLMs.'' 2024.
\bibitem{spinquant} Z.~Liu et al. ``SpinQuant: LLM Quantization with Learned Rotations.'' 2024.
\bibitem{quip} J.~Chee et al. ``QuIP / QuIP\#: 2-Bit Quantization with Incoherence Processing.'' 2023--2024.
\bibitem{qvla} ``QVLA: Per-Channel Action-Sensitivity Quantization for VLAs.'' (verify) 2026.
\bibitem{quantvla} ``QuantVLA: Scale-Calibrated PTQ for VLAs.'' arXiv:2602.20309, 2026.
\bibitem{omegaqvla} ``$\Omega$-QVLA: Robust Quantization for Vision-Language-Action Models via Composite Rotation and Per-step Scaling.'' 2026. Code: \texttt{github.com/UCMP13753/Omega-QVLA}.
\bibitem{dyqvla} ``DyQ-VLA: Temporal-Dynamic-Aware Quantization for Embodied Vision-Language-Action Models.'' arXiv:2603.07904, 2026.
\end{thebibliography}

\end{document}
