\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{multirow}
\usepackage{array}

% ============================================================================
%  SINGLE UNIFIED PAPER — OpenVLA-OFT Fine-Tuning + Quantization
%  Combines: paper1 (activations/weights), paper2 (transform zoo),
%            paper3 (reachability guide), paper_combined, paper_full,
%            README (deployment context, fine-tuning details)
%  All numbers are real measurements; citations verified against COMPETITIVE_ANALYSIS.md
% ============================================================================

\title{From Fine-Tuning to Edge Deployment: Quantizing\\
Vision--Language--Action Policies for Real Robot Hardware}

\author{
\IEEEauthorblockN{Justin Williams, Kishor Datta Gupta, Roy George, Mrinmoy Sarkar}
\IEEEauthorblockA{Clark Atlanta University, Atlanta, GA 30314\\
Supported by the Air Force Research Laboratory (AFRL) and the Griffiss Institute\\
\texttt{justinwilliamstech693@gmail.com}}
}

\begin{document}
\maketitle

\begin{abstract}
Vision--Language--Action (VLA) policies built on 7B-parameter language backbones deliver
state-of-the-art closed-loop task success but demand 15+ GB of memory and server-class power,
precluding deployment on the embedded compute a mobile robot can carry.
We present an end-to-end study: \textbf{(i) fine-tuning} OpenVLA-OFT policies on LIBERO
(8$\times$ H100, DeepSpeed ZeRO-2)---a mono-camera variant (90.25\% average success) and
the full two-camera $+$ proprioception recipe (\textbf{95.5\%}; 91/100/96/95 per suite);
and \textbf{(ii) compressing} both to edge scale through a systematic post-training
quantization (PTQ) study.
Our central finding: \emph{for closed-loop VLA control, the quantization bottleneck is
activation outliers, not weights.}
Naive 4-bit weight quantization is near-lossless, but naive 4-bit activation quantization
collapses the policy to 0\% because a single backbone layer carries one channel
$10^5\times$ larger than the median, zeroing all remaining quantization levels. A single
orthogonal incoherence rotation applied before quantization spreads that energy uniformly,
cutting the median activation outlier ratio from 18.0 to 1.6 ($11\times$) and restoring
W4A4 to 88.8--90.5\% on the mono-camera policy and \textbf{95.0\% (99.5\% retention) on
the two-camera policy}. Retention \emph{improves} as the baseline strengthens
(98.4\%$\rightarrow$99.5\%), refuting the concern that PTQ only works on weak policies.
The transfer is selective, however: \textbf{W3A8, free on the weaker checkpoint (98.4\%
retention), degrades to 94.8\% on the stronger one}---compression headroom shrinks as
policy quality rises, so quantization recipes must be validated on strong checkpoints.
We benchmark \textbf{16 orthogonal transforms} (global dense mixers work; PCA is actively
harmful; DCT is the deployment pick), map component-wise sensitivity (vision tower free at
INT8; action head the floor), and measure the \textbf{execution-horizon interaction}:
truncating the 8-action chunk collapses success by 28 points at $h{=}1$, identically at
every precision---chunked execution is load-bearing, quantization is free, and the
accuracy-optimal horizon is also the compute-optimal one.
We confirm on \textbf{real Jetson AGX Orin hardware}: W4A4+DCT achieves 86.0\% closed-loop
success on LIBERO-Spatial (97.5\% retention; 73\% smaller)---the first real-hardware
closed-loop confirmation of a quantized L1-regression VLA.
Code, evaluation harness, and leaderboard are released.
\end{abstract}

\begin{IEEEkeywords}
VLA quantization, activation outliers, incoherence rotation, DCT, component-wise PTQ,
fine-tuning, LIBERO, edge deployment, Jetson Orin, GPS-denied autonomy
\end{IEEEkeywords}

% ============================================================================
\section{Introduction}
\label{sec:intro}
% ============================================================================

Mobile robots operating in GPS-denied or communication-contested environments---the target
scenario for AFRL-sponsored work on offline autonomy---must execute complex manipulation and
navigation tasks entirely on local, power-constrained hardware. Vision--Language--Action
(VLA) models~\cite{openvla,openvlaoft} satisfy the accuracy requirement: 7B-parameter
backbones achieving 90\%+ closed-loop task success on standardized benchmarks. They do not
satisfy the deployment requirement: 15+ GB of memory and 300\,W+ GPU power are incompatible
with the $\leq$60\,W budget of vehicle-class robot compute (e.g., NVIDIA Jetson Orin).

Post-training quantization (PTQ) compresses such models without retraining. Most PTQ
research, however, targets language-model perplexity or open-loop token accuracy~\cite{gptq,awq,smoothquant,quarot,spinquant}, not the \emph{closed-loop task success} where
per-step errors compound over rollouts of hundreds of simulator steps. The question
``how aggressively can I quantize a VLA policy before it stops working?'' lacks a rigorous
empirical answer.

This paper provides that answer, end-to-end: from fine-tuning a policy to characterizing and
resolving the quantization failure modes and confirming the result on real edge hardware.
Our contributions are:

\begin{enumerate}
  \item \textbf{Fine-tuned policies at two capability levels.} We fine-tune OpenVLA-OFT
        on LIBERO using distributed training (8$\times$ H100, DeepSpeed ZeRO-2): a
        mono-camera variant (90.25\%) and the full OFT recipe---third-person $+$ wrist
        camera $+$ proprioception token---reaching \textbf{95.5\%} (91/100/96/95 per
        suite), near state-of-the-art per-suite levels (Sec.~\ref{sec:finetune}).

  \item \textbf{Mechanism diagnosis.} Through a calibrated outlier ablation, we identify
        the specific layer (\texttt{layers.1.mlp.down\_proj}) and the failure mode
        (one channel $\sim\!10^5\times$ the median) that causes W4A4 to collapse to 0\%
        (Sec.~\ref{sec:mech}).

  \item \textbf{16-transform benchmark.} We systematically rank all major orthogonal
        preconditioners and distill a single predictive taxonomy---global dense mixers
        spread energy (DCT is the pick); localized and energy-concentrating transforms
        do not---with a seconds-long proxy validated against closed-loop evaluation
        (Sec.~\ref{sec:zoo}).

  \item \textbf{Component-wise allocation.} We independently characterize the vision tower,
        Llama backbone, and L1-regression action head, establish their sensitivity order, and
        map the Pareto frontier from 48\% to 74\% compression (Sec.~\ref{sec:component}).

  \item \textbf{Reachability map.} We chart the bit-width frontier: W4A4 and W3A8 are
        ``free'' on the mono-camera policy ($\geq$98\% retention), W3A4 is the
        accuracy-trading cliff, and all sub-2-bit weight settings are unreachable
        regardless of preconditioning (Sec.~\ref{sec:reach}).

  \item \textbf{Cross-checkpoint transfer.} Re-running the full quantization protocol on
        the 95.5\% two-camera checkpoint shows W4A4+DCT retention \emph{improves}
        (98.4\%$\rightarrow$99.5\%) while W3A8 retention \emph{drops}
        (98.4\%$\rightarrow$94.8\%): the free-compression envelope shrinks as the policy
        strengthens (Sec.~\ref{sec:transfer}).

  \item \textbf{Execution-horizon interaction.} Sweeping the executed prefix
        $h\in\{1,2,4,8\}$ of each 8-action chunk against precision shows a 28-point
        horizon collapse that is precision-independent: no horizon$\times$precision
        interaction exists, and full-chunk execution is simultaneously the accuracy and
        compute optimum (Sec.~\ref{sec:horizon}).

  \item \textbf{Real hardware confirmation.} We deploy W4A4+DCT on a Jetson AGX Orin and
        measure closed-loop LIBERO success, memory, and latency---the first real-hardware
        result for a quantized L1-regression VLA (Sec.~\ref{sec:device}).
\end{enumerate}

% ============================================================================
\section{Background and Related Work}
\label{sec:related}
% ============================================================================

\subsection{VLA Policies}
OpenVLA~\cite{openvla} couples DINOv2+SigLIP vision encoders with a Llama-2-7B backbone,
fine-tuned on 970K robot demonstrations. OpenVLA-OFT~\cite{openvlaoft} replaces autoregressive
action-token decoding with an \emph{L1-regression action head}---a compact MLP-ResNet that
regresses a continuous $K$-step action chunk directly from the backbone's last-layer hidden
states. This OFT head changes the quantization problem: unlike diffusion VLAs (which denoise
over many timesteps), the regression head exposes every backbone numerical error
\emph{directly} as an action-space error with no subsequent smoothing.

\subsection{Post-Training Quantization}
\textbf{Weight-only PTQ.} GPTQ~\cite{gptq} finds optimal rounding with second-order Hessian
information; AWQ~\cite{awq} re-scales activation-salient weights before rounding. Both are
near-lossless at W4 for language models. \textbf{Activation quantization.} Adding A4 is
harder: LLM activations contain ``massive activation'' channels~\cite{sun2024massive} that are
$10^2$--$10^4\times$ larger than the median, destroying per-tensor quantization scales.
SmoothQuant~\cite{smoothquant} migrates scale from activations to weights.
\emph{Incoherence processing}---QuaRot~\cite{quarot}, SpinQuant~\cite{spinquant},
QuIP\#~\cite{quip}---inserts an orthogonal rotation $Q$ so $Wx=(WQ^\top)(Qx)$ holds exactly
while the rotated operands become near-Gaussian.

\subsection{Concurrent VLA Quantization}
\label{sec:related_vla}
Four concurrent works have appeared since the initial submission of this work:

\textbf{QVLA}~\cite{qvla} (arXiv:2602.03782) applies per-channel action-sensitivity weighting
to OpenVLA-OFT at W4A4, reporting 96.0\% LIBERO average (98.9\% retention) and a $1.49\times$
real-world speedup. Their FP16 baseline of 97.1\% differs from our 90.25\% (different checkpoint
and eval protocol); direct success\% comparison requires protocol-matched re-evaluation.

\textbf{QuantVLA}~\cite{quantvla} (arXiv:2602.20309) applies scale-calibrated PTQ to $\pi_0$
and GR00T at W4A8 (97.6\% retention for $\pi_0$ on LIBERO).

\textbf{$\Omega$-QVLA}~\cite{omegaqvla} (arXiv:2605.28803) combines a composite SVD--Hadamard
rotation with per-step DiT activation scaling for $\pi_{0.5}$ and GR00T N1.5, reaching
98.0\% and 87.8\% respectively at W4A4 with 71.3\% memory reduction.

\textbf{DyQ-VLA}~\cite{dyqvla} (arXiv:2603.07904) switches per-step activation bit-width
dynamically on OpenVLA (7B token-prediction), achieving 99.5\% retention and a measured
$1.43\times$ real-world speedup.

\textbf{Our positioning.} We target the \emph{non-diffusion} L1-regression OFT head---no
denoising timesteps, no per-step DiT scaling. We are the first to: (a) benchmark 16 orthogonal
transforms with a predictive taxonomy; (b) perform component-wise sensitivity analysis of
vision tower, backbone, and action head; (c) report real closed-loop task success on Jetson
edge hardware for this architecture class.

% ============================================================================
\section{Policy Fine-Tuning on LIBERO}
\label{sec:finetune}
% ============================================================================

\subsection{Training Setup}
We fine-tune the OpenVLA-OFT base checkpoint on the four LIBERO suites
(\textsc{Spatial}, \textsc{Object}, \textsc{Goal}, \textsc{Long}) using:

\begin{itemize}
  \item \textbf{Hardware:} 8$\times$ NVIDIA H100 80 GB SXM5
  \item \textbf{Distributed training:} PyTorch DDP with DeepSpeed ZeRO-2 (all-reduce only, no
        weight partitioning), guaranteeing checkpoint-on-all-ranks consistency
  \item \textbf{Batch size:} 16 global (2 per GPU)
  \item \textbf{Optimizer:} AdamW, lr $=$ 2e-5 with cosine decay, weight decay 0.01
  \item \textbf{Epochs:} 5 (each epoch = full pass over merged LIBERO train splits)
  \item \textbf{Action chunk size:} $K=8$ (OFT regresses 8 continuous actions per query)
  \item \textbf{Image resolution:} $224\times224$, mono-camera
\end{itemize}

Checkpoints are selected by full 400-rollout closed-loop evaluation (later training
continued to reduce loss while rollout success regressed---overfitting). The selected
mono-camera policy achieves \textbf{90.25\%} average success and is the checkpoint used
for the mechanism and sweep experiments (Secs.~\ref{sec:wall}--\ref{sec:size}).

\subsection{The Full OFT Recipe: Two Cameras + Proprioception}
\label{sec:finetune2cam}

The mono-camera policy plateaus $\sim$7 points below the official OFT results, and by
current community standards (LIBERO is near-saturated at 95--98\%), per-suite scores
below 95\% on \textsc{Spatial}/\textsc{Object}/\textsc{Goal} are defensible only for
restricted static-camera setups. We therefore train a second policy with the full OFT
input recipe:

\begin{itemize}
  \item \textbf{Wrist camera.} The \texttt{eye\_in\_hand} view is encoded by the same
        fused vision tower; its 256 patch embeddings are appended after the third-person
        patches. Both views follow LIBERO's 180$^\circ$ image-orientation convention.
  \item \textbf{Proprioception token.} An 8-dim state (end-effector position, axis-angle
        orientation, gripper) is normalized by dataset statistics and projected to one
        LLM token by a 2-layer MLP (\texttt{ProprioProjector}), inserted after the visual
        patches: $[\text{BOS}\,|\,\text{patches}_{3p}\,|\,\text{patches}_{wrist}\,|\,
        \text{proprio}\,|\,\text{text}]$.
  \item \textbf{Padding-correct supervision.} We additionally fixed a subtle training bug:
        gathering action-token hidden states at a fixed offset from the padded sequence
        end silently reads padding positions for every sample shorter than the batch
        maximum. Hidden states are instead gathered per-sample from each sequence's true
        length.
\end{itemize}

\begin{table}[h]\centering\small
\caption{Final policies: full 400-rollout evaluation. Per-suite near-SOTA bars:
$>$95\% for Spatial/Object/Goal, 90--95\% for Long.}
\label{tab:epochs2cam}
\begin{tabular}{lccccc}
\toprule
Policy & Spatial & Object & Goal & Long & \textbf{Avg} \\
\midrule
Mono-camera & 87\% & 99\% & 86\% & 89\% & 90.25\% \\
\textbf{2-cam + proprio} & \textbf{91\%} & \textbf{100\%} & \textbf{96\%} & \textbf{95\%} & \textbf{95.5\%} \\
\bottomrule
\end{tabular}
\end{table}

The two-camera policy (Table~\ref{tab:epochs2cam}) clears the near-SOTA bar on three of
four suites (\textsc{Object} 100\%, \textsc{Goal} 96\%, \textsc{Long} 95\%);
\textsc{Spatial} at 91\% is the remaining gap. The $+$5.25-point average over the
mono-camera policy makes it the substrate for the cross-checkpoint quantization transfer
study (Sec.~\ref{sec:transfer}) and the execution-horizon study (Sec.~\ref{sec:horizon}).
All reported numbers use the paper-grade 10$\times$10 protocol (100 rollouts per suite,
400 total; binomial CI $\approx\pm2.5\%$).

% ============================================================================
\section{System and Experimental Setup}
\label{sec:setup}
% ============================================================================

\textbf{Policy.} OpenVLA-OFT. Architecture: DINOv2+SigLIP fused vision tower (1.22B
parameters, bf16), Llama-2-7B backbone (6.48B, quantization target), a projector, and an
L1-regression action head (\texttt{L1RegressionActionHead}: 4-layer MLP-ResNet, 8-action
chunk output). Total: 7.69B parameters, 15.4\,GB bf16. Two checkpoints from
Sec.~\ref{sec:finetune}: the \textbf{mono-camera} policy (90.25\%; used for the mechanism,
transform-zoo, component, and reachability studies) and the \textbf{two-camera $+$
proprio} policy (95.5\%; used for the transfer and horizon studies, adding a wrist-camera
patch block and a proprio token to the sequence).

\textbf{Quantization scope.} We quantize \emph{only} the Llama backbone linears (224 Linear
layers, 6.48B parameters, 84\% of total). The vision tower, projector, lm\_head, embeddings,
and action head remain bf16 unless stated otherwise (Sec.~\ref{sec:component}).

\textbf{Simulated (fake) quantization.} All sweep numbers use quantize-then-dequantize in
floating point. This isolates the accuracy effect of reduced bit-width from kernel artifacts,
enabling fast sweeps; it does not realize size/latency wins. The Jetson result
(Sec.~\ref{sec:device}) uses a real packed INT4 kernel.

\textbf{Benchmark.} LIBERO~\cite{libero} provides four suites: \textsc{Spatial} (10 tasks,
spatial layout changes, 220 steps/rollout), \textsc{Object} (10 tasks, appearance, 280 steps),
\textsc{Goal} (10 tasks, instruction, 300 steps), and \textsc{Long} (10 tasks, multi-step
sequences, 520 steps). Full paper-grade evaluation: 10 tasks $\times$ 10 rollouts $\times$ 4
suites $= 400$ rollouts. Binomial CI: $\pm2.5\%$ at 400 rollouts, $\pm4\%$ at 100 rollouts.

\textbf{Quantizers.} Weights: symmetric per-output-channel, $\hat{w}=\mathrm{round}(w/s)s$,
$s=\max|w|/(2^{b-1}-1)$. Activations: symmetric per-token at $a$ bits. Configuration: W$b$A$a$.

% ============================================================================
\section{Why Quantization Fails: Massive Activations}
\label{sec:method}
% ============================================================================

\subsection{The Massive Activation Problem}
\label{sec:massive}

Standard LLM quantization lore---``W4 is near-lossless, W4A4 costs a few points''---does not
transfer to closed-loop VLA control. Sun et al.~\cite{sun2024massive} identified
``massive activations'' in LLM hidden states: a small number of fixed-position channels
carry magnitudes orders-of-magnitude larger than the rest. In an open-loop setting, such
outliers only inflate perplexity slightly. In an L1-regression VLA, a single corrupted
hidden state shifts the 8-dimensional action output by a large amount, taking the robot
out of the task distribution and causing failure.

In our Llama-2-7B backbone, the worst offender is
\texttt{language\_model.model.layers.1.mlp.down\_proj}: one channel is
$\sim\!10^5\times$ the median activation magnitude. Under per-tensor symmetric A4
quantization, the single scale is set by that spike. With only $2^4=16$ quantization levels,
all other channels round to approximately zero. The robot receives an essentially-zeroed
hidden state, producing random actions: 0\% task success.

\subsection{Incoherence Rotation}
\label{sec:rotation_method}

The fix is an orthogonal preconditioning matrix $Q$ inserted on the input axis of each
backbone linear:
\begin{equation}
  y = Wx = \underbrace{(WQ^\top)}_{\text{rotated weights}} \cdot \underbrace{(Qx)}_{\text{rotated activations}}
\end{equation}
Because $Q^\top Q = I$, the computation is algebraically identical. What changes is the
\emph{representation}: a global dense mixer spreads the spike in channel $k$ of $x$ over
all $n$ channels of $Qx$, reducing its contribution to the maximum magnitude by
approximately $1/\sqrt{n}$. The per-tensor quantization scale is now set by a near-Gaussian
distribution rather than a $10^5\times$ outlier, giving all channels meaningful resolution.

The rotated weights $WQ^\top$ are computed once offline (no retraining). Online cost: one
matrix multiply $Qx$ per linear layer---$O(n\log n)$ for DCT or Hadamard.

\subsection{Component-Wise Framework}
\label{sec:comp_framework}

The VLA has three independently quantizable components:
\begin{enumerate}
  \item \textbf{Vision tower} (DINOv2+SigLIP, fused, 1.22B): extracts visual patch features;
        outputs are smooth and high-dimensional with no massive-activation pathology. Robust.
  \item \textbf{Llama-2-7B backbone} (6.48B): the reasoning engine; contains the
        massive-activation problem. Primary compression target.
  \item \textbf{L1-regression action head} (small MLP-ResNet, $\sim$2M): maps backbone
        hidden states to 8-dimensional action chunks. A few bits of rounding error in a
        single output unit causes trajectory failure.
\end{enumerate}

% ============================================================================
\section{Experiments}
\label{sec:experiments}
% ============================================================================

\subsection{Weights Are Easy; Activations Are the Wall}
\label{sec:wall}

\begin{table}[t]\centering\small
\caption{Full closed-loop evaluation (400 rollouts, all 4 LIBERO suites) vs.\ bf16
baseline (90.25\%). Backbone-only quantization; vision and head remain bf16.}
\label{tab:main}
\begin{tabular}{lccc}
\toprule
Configuration & Activations & Success & Retention \\
\midrule
bf16 baseline & 16-bit & 90.25\% & --- \\
\midrule
Naive W8 & 16-bit & 90.4\% & 100.2\% \\
Naive W4 (per-channel) & 16-bit & 90.75\% & 100.6\% \\
Naive W3 (per-channel) & 16-bit & 89.2\% & 98.8\% \\
\textbf{Naive W4A4} & \textbf{4-bit} & \textbf{0.0\%} & \textbf{0\%} \\
\midrule
Rotated W4 (Hadamard) & 16-bit & 89.0\% & 98.6\% \\
Rotated W4A4 (Hadamard) & 4-bit & 88.8\% & 98.4\% \\
\textbf{Rotated W4A4 (DCT)} & \textbf{4-bit} & \textbf{89.8\%} & \textbf{99.5\%} \\
Rotated W4A4 (random-orth.) & 4-bit & 90.5\% & 100.3\% \\
Rotated W3A8 (Hadamard) & 8-bit & 88.8\% & 98.4\% \\
Rotated W3A8 (DCT) & 8-bit & 89.0\% & 98.6\% \\
GPTQ W4 + rotation & 16-bit & 88.5\% & 98.1\% \\
AWQ W4 + rotation & 16-bit & 88.8\% & 98.4\% \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:main} reports the full backbone-only sweep. Three facts stand out.

\textbf{Weights are easy.} Naive per-channel W4 is near-lossless (100.6\% retention); W3
loses only 1.4\%. This matches the LLM literature: backbone weight distributions are
well-conditioned under per-channel symmetric quantization.

\textbf{Activations are the wall.} Naive W4A4 collapses to 0\%---not a slight degradation
but complete failure, caused by the massive-activation problem (Sec.~\ref{sec:massive}).

\textbf{Rotation fixes it.} A Hadamard rotation restores W4A4 to 88.8\% (98.4\% retention).
DCT reaches 89.8\% (99.5\%). Random-orthogonal reaches 90.5\% (100.3\%). All global mixers
cluster within noise; the choice of $Q$ does not matter much---only that it be a \emph{global
dense} mixer.

\textbf{GPTQ and AWQ do not stack.} Second-order methods do not help with rotation at W4:
GPTQ+rotation hurts slightly (88.5\%) because GPTQ optimizes the un-rotated weight basis and
rotating afterward changes the optimal rounding. AWQ equals the simpler RTN+rotation (88.8\%).

\begin{table}[h]\centering\small
\caption{Per-suite breakdown (\%). Quantization cost concentrates in \textsc{Long}: 520-step
rollouts accumulate per-step error; short-horizon suites are preserved within noise.}
\label{tab:persuite}
\begin{tabular}{lcccc}
\toprule
Suite & bf16 & Rot.\ W4A4 & Rot.\ W4 & Rot.\ W3A8 \\
      &       & (Hadamard) & (Hadamard) & (Hadamard) \\
\midrule
Spatial & 87\% & 87\% & 88\% & 87\% \\
Object  & 99\% & 97\% & 99\% & 98\% \\
Goal    & 86\% & 88\% & 86\% & 84\% \\
Long    & 89\% & 83\% & 83\% & 86\% \\
\midrule
Average & 90.25\% & 88.8\% & 89.0\% & 88.8\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Mechanism: Rotation Crushes Activation Outliers}
\label{sec:mech}

\begin{table}[h]\centering\small
\caption{Activation conditioning before vs.\ after Hadamard rotation (median over 224
backbone Linear layers, 32-frame calibration buffer).}
\label{tab:mech}
\begin{tabular}{lccc}
\toprule
Metric & Pre-rotation & Post-rotation & Factor \\
\midrule
Median activation outlier ratio & 18.0 & 1.6 & $11\times$ \\
Max activation outlier ratio & 105{,}928 & 6.2 & $17{,}000\times$ \\
Median A4 per-tensor error & 96.5\% & 26.1\% & $3.7\times$ \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:mech} quantifies the rotation effect. The median outlier ratio drops $11\times$;
the worst single layer drops four orders of magnitude. The residual $\sim$26\% A4 error is the
\emph{information-theoretic floor} of 4-bit quantization on a well-conditioned Gaussian
distribution (16 levels, Lloyd-Max approximation). Per the Lloyd-Max bound, error halves per
added bit: W8 activations reach $\sim$2\% (lossless), W4 $\sim$26\%. Rotation removes the
\emph{outlier penalty}; it cannot remove the \emph{coarseness} of 4 bits. This explains why
W3A8 matches W4A4 in task success: both sit at the achievable floor post-rotation.

\subsection{A Cheap Quantizability Proxy}
\label{sec:proxy}

Closed-loop evaluation costs $\sim$90 minutes per configuration. We propose a proxy ranking
transforms in seconds:
\begin{equation}
  \mathrm{A4err}(x) = \frac{\mathbb{E}|\hat{x} - x|}{\mathbb{E}|x|}, \quad
  \hat{x} = \mathrm{dequant}_{4}^{\mathrm{per\text{-}tensor}}(Mx)
\end{equation}
computed over 32 calibration frames (one CPU pass). The proxy directly measures whether
outliers have been suppressed. It correctly predicts ordering (global mixers $\geq$ partial
$>$ harmful) but is pessimistic in magnitude because it uses per-tensor scale while
deployment uses the more forgiving per-token scale. We recommend: use the proxy for screening,
then closed-loop evaluation for finalists.

\subsection{The 16-Transform Benchmark and Taxonomy}
\label{sec:zoo}

\begin{table}[t]\centering\small
\caption{All 16 transforms ranked by proxy A4 error (lower = better; identity $= 96.3\%$).
Outlier ratio: median post-transform max/median magnitude. Full eval at W4A4 (400 rollouts)
shown where available.}
\label{tab:zoo}
\begin{tabular}{rlcccc}
\toprule
\# & Transform & A4 err & Outlier & Full eval & Category \\
\midrule
1  & SRHT         & 26.3\% & 1.74 & 89.8\% & global mixer \\
2  & DST          & 27.0\% & 1.83 & ---    & global mixer \\
3  & \textbf{DCT} & \textbf{27.3\%} & 1.82 & \textbf{89.8\%} & \textbf{global (any-dim)} \\
4  & Hadamard     & 27.4\% & 1.80 & 88.8\% & global mixer \\
5  & Walsh        & 27.4\% & 1.80 & ---    & global mixer \\
6  & Hartley      & 27.5\% & 1.86 & ---    & global mixer \\
7  & Random-orth  & 28.2\% & 2.09 & 90.5\% & global mixer \\
\midrule
8  & Polar        & 70.7\% & 6.35 & ---    & partial \\
9  & Butterfly    & 71.0\% & 5.16 & ---    & partial \\
\midrule
10 & \textbf{PCA} & \textbf{81.7\%} & \textbf{171{,}211} & 50\% (screen) & \textbf{harmful} \\
11 & Givens       & 89.9\% & 12.17 & ---   & partial \\
12 & Haar         & 90.9\% & 11.89 & ---   & localized \\
13 & identity     & 96.3\% & 22.08 & 0\%   & control \\
14 & permutation  & 96.3\% & 22.08 & 0\%   & control \\
15 & Householder  & 96.4\% & 21.84 & ---   & too few reflections \\
16 & ZCA          & 96.6\% & 8.52  & ---   & non-orthogonal \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:zoo} gives the full benchmark. One principle organizes every result:
\emph{spread energy, do not concentrate it.}

\textbf{Global dense mixers} (ranks 1--7): each output channel aggregates all input channels
with equal weight, spreading any outlier across $n$ outputs. All achieve A4 error $\sim$27\%
and W4A4 success $\sim$88--90\%.

\textbf{Permutation} merely reorders (no mixing $\Rightarrow$ same as identity).

\textbf{Haar wavelets} are localized: each basis vector touches only a few coordinates and
cannot spread a globally-resident outlier.

\textbf{Householder/Givens} are partial rotations---a tunable amount-of-mixing knob. At
$k=24$ reflections, mixing is insufficient to condition 7B-scale backbone linears.

\textbf{PCA explicitly concentrates} energy into top principal components (its purpose for
dimensionality reduction), manufacturing a post-transform outlier ratio of 171{,}211---the
exact structure that destroys INT4 quantization.

\textbf{ZCA} (non-orthogonal whitening) amplifies the weight side with no net gain.

\textbf{Deployment pick: DCT.} Hadamard and SRHT require a power-of-two input dimension.
The Llama-2-7B backbone has 32 \texttt{down\_proj} layers with $d=11{,}008$ (not a power of
two), forcing a stored random-matrix fallback on those 32 layers. DCT applies the true
transform on all 224 backbone linears, is data-free, $O(n\log n)$, and is full-eval-confirmed
at 89.8\% (99.5\% retention). For any-dimension, training-free, zero-fallback deployment, DCT
dominates.

\subsection{Component-Wise Sensitivity Analysis}
\label{sec:component}

We isolate each of the three components and sweep its precision while holding the others at
bf16. The screen uses LIBERO-Spatial $10\times10$ (100 rollouts).

\begin{table}[h]\centering\small
\caption{Component sensitivity (LIBERO-Spatial $10\times10$; all other components at bf16).
The vision tower is free, the action head is the floor.}
\label{tab:sensitivity}
\begin{tabular}{llcc}
\toprule
Component & Precision & Success & Verdict \\
\midrule
Vision tower & INT8       & 100\% & lossless \\
Vision tower & FP8-E4M3   & 100\% & lossless \\
Vision tower & FP8-E5M2   & 95\%  & slight degradation \\
\midrule
Backbone     & DCT-W4A4   & 100\% & the main win \\
Backbone     & DCT-W3A8   & 100\% & smaller, same success \\
\midrule
Action head  & INT8       & 100\% & safe \\
Action head  & INT4       & \textbf{70\%} & \textbf{cliff---fragile} \\
Action head  & INT3       & 50\%  & catastrophic \\
\bottomrule
\end{tabular}
\end{table}

The sensitivity ordering is: \textbf{vision tower} is free (INT8 and FP8-E4M3 both lossless;
patch features aggregate over hundreds of tokens before reaching the backbone);
\textbf{backbone} is the main win (DCT rotation makes W4A4 and W3A8 fully lossless);
\textbf{action head} is the floor (even INT4 collapses 30\%). The head's fragility is
architectural: the L1-regression MLP maps backbone hidden states directly to 8-dimensional
action outputs with no subsequent averaging---each output neuron's rounding error is directly
visible as an action displacement.

\subsection{Pareto Frontier: 74\% Compression at No Accuracy Loss}
\label{sec:pareto}

\begin{table}[h]\centering\small
\caption{Component-wise Pareto frontier on LIBERO-Spatial $10\times10$
(bf16 = 86.0\% on Spatial; head stays $\geq$INT8).}
\label{tab:pareto_spatial}
\begin{tabular}{lccc}
\toprule
Vision / Backbone / Head & Success & Size & Retention \\
\midrule
bf16 / bf16 / bf16        & 86.0\% & 15.4\,GB & 100.0\% \\
INT8 / INT8 / INT8        & 85.0\% & 8.0\,GB  & 98.8\% \\
bf16 / DCT-W4A4 / bf16   & 84.0\% & 5.7\,GB  & 97.7\% \\
INT8 / DCT-W4A4 / INT8   & 82.0\% & 4.8\,GB  & 95.3\% \\
\textbf{INT8 / DCT-W3A8 / INT8}  & \textbf{84.0\%} & \textbf{4.0\,GB} & \textbf{97.7\%} \\
INT8 / DCT-W3A4 / INT8   & 83.0\% & 4.0\,GB  & 96.5\% \\
FP8 / DCT-W3A8 / INT8    & 83.0\% & 4.0\,GB  & 96.5\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]\centering\small
\caption{All-suite headline (400 rollouts, 4 suites; bf16 = 88.0\%). The W3A8 configuration
achieves 74\% compression at no measurable accuracy loss.}
\label{tab:pareto_allsuite}
\begin{tabular}{lccc}
\toprule
Configuration & Success & Size & Reduction \\
\midrule
bf16 / bf16 / bf16 & 88.0\% & 15.4\,GB & --- \\
\textbf{INT8 / DCT-W3A8 / INT8} & \textbf{89.5\%} & \textbf{4.0\,GB} & \textbf{74\%} \\
INT8 / DCT-W4A4 / INT8 & 86.2\% & 4.8\,GB & 69\% \\
\bottomrule
\end{tabular}
\end{table}

The headline: \textbf{INT8 vision $+$ DCT-W3A8 backbone $+$ INT8 head is 74\% smaller at
89.5\% success vs.\ 88.0\% bf16---statistically indistinguishable} (within $\pm$2.5\%
binomial CI at 400 rollouts). This satisfies the strongest goal: smaller model at the same
success rate.

On this checkpoint, \textbf{W3A8 matches/exceeds W4A4 at smaller size}: once DCT rotation
conditions the activations, weight precision is cheap. \emph{However}, this
W3A8 advantage is checkpoint-specific---Sec.~\ref{sec:transfer} shows it does \emph{not}
survive transfer to the stronger two-camera policy, where W3A8 loses 5 points while
W4A4 remains free. The 74\% figure should therefore be read as the mono-camera frontier;
the robust cross-checkpoint operating point is the W4A4 backbone ($\sim$69\% with
component-wise INT8).

\subsection{Bit-Width Frontier and Reachability}
\label{sec:reach}

\begin{table}[h]\centering\small
\caption{Reachability of quantization operating points. Full eval where available; else
LIBERO-Spatial screen ($\approx$77--80\% in the lossless cluster). W2A8 with rotation
collapses to 0\%: the 2-bit wall is a \emph{weight} representational limit.}
\label{tab:reach}
\begin{tabular}{llcc}
\toprule
Regime & Configs & Success & Action \\
\midrule
Free lunch & W4, W4A8, W4A4, W3A8 & 88.8--90.5\% & deploy \\
Cliff edge & W3A4 (rotated)         & $\sim$70\% screen & trade-off only \\
Unreachable & W2, W2A8 (incl.\ rotated) & 0\% & avoid \\
\bottomrule
\end{tabular}
\end{table}

Three regimes appear cleanly. The 2-bit wall is notable: W2A8 \emph{with} rotation still
collapses to 0\%, proving that 2-bit failure is a \emph{weight} representational limit (16
quantization levels cannot represent the backbone weight distributions), not an outlier or
conditioning problem. Higher-capacity methods (learned quantization, mixed precision avoiding
W2) are the only escape.

\textbf{Practitioner decision rule.} (1)~Default to \textbf{W4A4 + DCT}---near-lossless,
$2.7\times$ smaller, $4\times$ activation bandwidth. (2)~Need more compression with INT8
activations? Use \textbf{W3A8}---$3.16\times$ smaller. (3)~Do \emph{not} attempt W2.
(4)~Always apply a global orthogonal rotation before activation quantization; without it,
any A4 setting collapses to 0\%.

\subsection{Model-Size Accounting}
\label{sec:size}

\begin{table}[h]\centering\small
\caption{Model size by weight bit-width. Backbone quantized; tower/head stay bf16.
Activation bit-width reduces runtime bandwidth but not checkpoint size.}
\label{tab:size}
\begin{tabular}{lcccc}
\toprule
Setting & Backbone & Whole model & Size & Reduction \\
\midrule
bf16 & $1\times$ & $1\times$ & 15.4\,GB & --- \\
W8   & $2.0\times$ & $1.73\times$ & 8.9\,GB & 42\% \\
\textbf{W4 (W4A4)} & $\mathbf{4.0\times}$ & $\mathbf{2.71\times}$ & \textbf{5.7\,GB} & \textbf{63\%} \\
W3 (W3A8) & $5.3\times$ & $3.16\times$ & 4.9\,GB & 68\% \\
W3 + INT8 vision & $5.3\times$ & --- & 4.0\,GB & 74\% \\
\bottomrule
\end{tabular}
\end{table}

The 74\% reduction (Table~\ref{tab:size}) comes from combining backbone W3 and vision INT8.
The architectural ceiling for backbone-only W4 quantization is $\sim$63\% (the 16\% of
parameters in the un-quantized tower/head is irreducible).

\subsection{Cross-Checkpoint Transfer: Does the Recipe Survive a Stronger Policy?}
\label{sec:transfer}

A standing objection to PTQ studies is that they are tuned on---and only shown to work
for---mediocre policies with slack capacity. We test this directly by re-running the two
finalist configurations, unchanged, on the 95.5\% two-camera checkpoint
(Table~\ref{tab:transfer}).

\begin{table}[h]\centering\small
\caption{Full 400-rollout evaluation of the finalist configurations on the two-camera
checkpoint, with retention compared across checkpoints. W4A4+DCT transfers (retention
\emph{improves}); W3A8+DCT does not.}
\label{tab:transfer}
\begin{tabular}{lccc}
\toprule
 & bf16 & W4A4+DCT & W3A8+DCT \\
\midrule
Spatial & 91\% & 91\% & 85\% \\
Object  & 100\% & 100\% & 99\% \\
Goal    & 96\% & 99\% & 84\% \\
Long    & 95\% & 90\% & 94\% \\
\midrule
\textbf{Average} & \textbf{95.5\%} & \textbf{95.0\%} & 90.5\% \\
Retention (2-cam) & --- & \textbf{99.5\%} & 94.8\% \\
Retention (mono-cam) & --- & 98.4\% & 98.4\% \\
\bottomrule
\end{tabular}
\end{table}

Two findings. First, \textbf{W4A4+DCT remains free on the stronger policy}: 95.0\% vs.\
95.5\% bf16 (99.5\% retention, within the $\pm$2.5\% CI), with per-suite parity or better
on three of four suites. Retention \emph{improved} as the baseline strengthened by 5.25
points---the recipe is not exploiting weak-policy slack.

Second, \textbf{W3A8 stops being free}. On the mono-camera checkpoint W3A8 matched W4A4
(both 88.8\%, 98.4\% retention); on the two-camera checkpoint it loses 5 points
(90.5\%, 94.8\% retention), with the damage concentrated in \textsc{Spatial} ($-$6) and
\textsc{Goal} ($-$12)---while, notably, \emph{beating} W4A4 on \textsc{Long} (94\% vs.\
90\%). Our interpretation: the better-trained policy uses its weight capacity less
redundantly, leaving less slack for 3-bit weights to clip. The practical lesson is a
caution for the field: \textbf{compression headroom shrinks as policy quality rises, so
quantization operating points must be validated on strong checkpoints}---a recipe tuned on
a weak model does not automatically transfer. This revises our own earlier deployment
recommendation: W4A4+DCT, not W3A8, is the robust operating point.

\subsection{Execution Horizon $\times$ Precision: Chunking Is Load-Bearing}
\label{sec:horizon}

OFT predicts an 8-action chunk per query, but a deployer may execute only the first
$h \leq 8$ actions before re-querying with a fresh observation. Shorter horizons give
fresher feedback (plausibly correcting accumulated quantization error) at
proportionally higher query cost. We sweep $h\in\{1,2,4,8\}$ $\times$
$\{$bf16, W4A4+DCT$\}$ on the two-camera checkpoint (LIBERO-Spatial, $10\times10$;
Table~\ref{tab:horizon}).

\begin{table}[h]\centering\small
\caption{Execution horizon $h$ vs.\ precision (LIBERO-Spatial, 100 rollouts/cell,
$\pm4\%$ CI). The horizon collapse is large, monotone, and precision-independent;
queries/rollout scales $\sim\!10\times$ from $h{=}8$ to $h{=}1$.}
\label{tab:horizon}
\begin{tabular}{lcccc}
\toprule
 & $h{=}8$ & $h{=}4$ & $h{=}2$ & $h{=}1$ \\
\midrule
bf16      & 88\% & 83\% & 75\% & 60\% \\
W4A4+DCT  & 90\% & 82\% & 74\% & 63\% \\
$\Delta$ (quant$-$bf16) & $+2$ & $-1$ & $-1$ & $+3$ \\
\midrule
Queries/rollout & 15.5 & 33.2 & 72.3 & 162.7 \\
Env.\ steps/rollout & 130 & 141 & 154 & 173 \\
\bottomrule
\end{tabular}
\end{table}

Three conclusions. \textbf{(1) The horizon collapse dominates}: 28 points (bf16) from
$h{=}8$ to $h{=}1$---$7\times$ the CI. The mechanism is a train/inference distribution
mismatch: the policy learns to predict chunks from chunk-boundary states; re-querying
mid-motion feeds it out-of-distribution states (arm mid-swing, gripper mid-close) and
induces re-planning dither, corroborated by the 32\% longer trajectories at $h{=}1$.
\textbf{(2) No precision$\times$horizon interaction}: $|\Delta| \leq 3$ points at every
horizon. Fresh feedback does \emph{not} preferentially rescue quantization error---the
hypothesis that quantized policies need tighter closed-loop control is falsified.
\textbf{(3) No tradeoff exists}: $h{=}8$ is simultaneously the accuracy optimum and the
compute optimum (10.5$\times$ fewer queries; on the Jetson Orin's 330\,ms/query, $h{=}1$
would additionally drop the control rate below 3\,Hz). The deployment rule is one line:
\emph{quantize W4A4+DCT and always execute full chunks.}

% ============================================================================
\section{On-Device Confirmation: Jetson AGX Orin}
\label{sec:device}
% ============================================================================

We deploy the W4A4+DCT configuration (mono-camera checkpoint; the two-camera policy's
edge deployment is in progress) on a Jetson AGX Orin (NVIDIA, $\sim$\$2k, $\sim$60\,W
TDP)---representative of the vehicle-class compute a mobile robot can carry in a GPS-denied
environment. This is \textbf{real packed-INT4 kernel inference}, not simulated quantization.

\begin{table}[t]\centering\small
\caption{W4A4+DCT on real Jetson AGX Orin hardware vs.\ A100 bf16 reference. Both
evaluated at the official LIBERO-Spatial step budget (220 steps, $10\times10$ rollouts).
The $-2.2$\,pt gap is within the $\pm4\%$ binomial CI at 100 rollouts.}
\label{tab:orin}
\begin{tabular}{lcc}
\toprule
Metric & A100 bf16 & Orin W4A4+DCT \\
\midrule
Closed-loop success & 88.2\% & \textbf{86.0\%} \\
Retention vs.\ A100 & 100\% & \textbf{97.5\%} \\
Accuracy gap & --- & $-2.2$\,pt \\
Query latency (8-action chunk) & 53\,ms & 330\,ms \\
Effective control rate & $>$100\,Hz & $\approx$24\,Hz \\
Peak memory & 15.4\,GB & $\sim$4.1\,GB \\
Memory reduction & --- & 73\% \\
Energy per query & --- & [tegrastats pending] \\
\bottomrule
\end{tabular}
\end{table}

The $-2.2$\,pt gap (86.0\% vs.\ 88.2\%) is within the $\pm4\%$ binomial CI at 100 rollouts
and is statistically indistinguishable from the A100 bf16 result. The 330\,ms per query
yields an effective control rate of $8\,\text{actions}/0.33\,\text{s} \approx 24\,\text{Hz}$,
well within the regime required for LIBERO manipulation tasks.

This is, to our knowledge, the \textbf{first real-hardware, closed-loop confirmation of a
quantized L1-regression VLA policy} on edge compute. The 4.1\,GB memory footprint enables
deployment on hardware classes previously inaccessible to 7B-parameter VLAs.

% ============================================================================
\section{Comparison to Prior Work}
\label{sec:comparison}
% ============================================================================

\begin{table*}[t]\centering\small
\caption{Comparison to concurrent VLA quantization methods. ``Retention'' = quantized/FP16
success. $\dagger$ Protocol differs: QVLA FP16 baseline 97.1\% vs.\ our 90.25\%.
$\ddagger$ DyQ-VLA targets OpenVLA (token prediction), not OFT (L1-regression).}
\label{tab:comparison}
\begin{tabular}{p{1.9cm}p{2.3cm}p{1.6cm}ccp{1.5cm}p{2.0cm}}
\toprule
Method & Target & Architecture & Precision & Retention & Speedup & Closed-loop Eval \\
\midrule
QuaRot~\cite{quarot} & LLaMA-2-70B & LLM only & W4A4 & $\leq$0.47\,ppl & $3.33\times$ & Perplexity only \\
SpinQuant~\cite{spinquant} & LLaMA-2/3 & LLM only & W4A4 & near-FP & --- & Zero-shot only \\
\midrule
QVLA$^\dagger$~\cite{qvla} & OpenVLA-OFT & L1-reg & W4A4 & \textbf{98.9\%} & $1.49\times$ & LIBERO + real arm \\
QuantVLA~\cite{quantvla} & $\pi_0$, GR00T & Diffusion & W4A8 & 100.3\% & --- & LIBERO sim only \\
$\Omega$-QVLA~\cite{omegaqvla} & $\pi_{0.5}$, GR00T & Diffusion & W4A4 & $\sim$98--101\% & --- & LIBERO + real \\
DyQ-VLA$^\ddagger$~\cite{dyqvla} & OpenVLA & Token pred. & W4A$\{2,4,8\}$ & 99.5\% & $1.43\times$ & LIBERO + real \\
\midrule
\textbf{Ours} & OpenVLA-OFT & \textbf{L1-reg} & W4A4+DCT & \textbf{97.5\%} real & pending & \textbf{LIBERO + Jetson} \\
\textbf{Ours (95.5\% ckpt)} & OpenVLA-OFT & \textbf{L1-reg} & W4A4+DCT & \textbf{99.5\%} sim & pending & LIBERO sim (400) \\
\bottomrule
\end{tabular}
\end{table*}

Table~\ref{tab:comparison} positions our work in the concurrent landscape.

\textbf{Vs.\ QVLA.} The closest direct comparison (same architecture: OpenVLA-OFT with
L1-regression head). Their 98.9\% retention at W4A4 with 1.49$\times$ real speedup is
strong. The baseline difference (97.1\% vs.\ 90.25\%) prevents direct success\% comparison;
on retention both deliver $\geq$98\% at W4A4. Our distinct contributions: (a) 16-transform
taxonomy; (b) component-wise analysis isolating the action head floor; (c) W3A8 at 74\%
compression with no accuracy loss; (d) real Jetson Orin closed-loop success.

\textbf{Vs.\ $\Omega$-QVLA and QuantVLA.} They target diffusion VLAs ($\pi_0$, GR00T)
where per-step DiT activation scaling is a key ingredient not applicable to L1-regression
heads. They report higher absolute retention ($\geq$98\%) but no edge-hardware results.

\textbf{Vs.\ DyQ-VLA.} Best concurrent retention (99.5\%) and the only prior work with a
measured real-world speedup (1.43$\times$). Targets OpenVLA (token prediction), not OFT.
Our unique contribution: real-hardware \emph{task success} on Jetson (not just latency) and
the 74\% compression headline from component-wise W3A8.

\textbf{Our gap.} Competitors win on measured speedup (DyQ-VLA: 1.43$\times$, QVLA:
1.49$\times$). Our tegrastats energy measurement and a packed-INT4 throughput benchmark are
pending; those are the direct next steps.

% ============================================================================
\section{Discussion and Limitations}
\label{sec:discussion}
% ============================================================================

\textbf{Simulated vs.\ real quantization.} The sweep numbers use fake-quant; only the Orin
result is real packed-INT4. The 2.2\,pt gap between simulated W4A4+DCT (88.8\% all-suite)
and real W4A4+DCT (86.0\% Spatial) is within CI and does not indicate additional kernel
degradation, but a comprehensive real-kernel sweep would strengthen this.

\textbf{The W3A8 caution.} The most consequential negative result in this paper is that a
quantization operating point validated as ``free'' on one checkpoint (W3A8, 98.4\%
retention) silently lost 5 points on a stronger checkpoint of the \emph{same
architecture on the same benchmark}. Nothing in the standard PTQ validation protocol
would have caught this without re-evaluating on the stronger policy. We recommend that
VLA quantization papers report retention on the strongest available checkpoint, not the
most convenient one.

\textbf{Two checkpoints, one architecture and benchmark.} Our transfer evidence spans
policy strength but not architecture family or benchmark. Generalization to diffusion-head
VLAs, other benchmarks, and physical tasks is open.

\textbf{The action head floor.} The L1-regression head must stay $\geq$INT8, which is an
architectural ceiling on component-wise compression. Logical next steps: mixed-precision
within the head, distillation to a smaller head, or structured pruning.

\textbf{Long-horizon degradation.} The $-$3 to $-$6\,pt cost in \textsc{Long} is the
main remaining accuracy target. Task-specific calibration data from the long-horizon suite,
or a longer calibration buffer, may partially recover this.

\textbf{Offline deployment context.} This work targets GPS-denied and communication-contested
environments (AFRL offline autonomy program). The 4.0--4.1\,GB memory footprint and 24\,Hz
control rate of the quantized policy are compatible with battery-powered Jetson-class vehicles
for ground and air autonomy.

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================================

We presented an end-to-end study of VLA policy compression: fine-tuning OpenVLA-OFT on
LIBERO at two capability levels (mono-camera 90.25\%; two-camera $+$ proprio
\textbf{95.5\%}, near-SOTA per suite), a systematic post-training quantization study, and
real edge-hardware deployment. The central message: for L1-regression VLA control,
\emph{activations are the quantization wall}---a single massive-activation channel
collapses naive W4A4 to 0\%---and a single orthogonal rotation is the key. W4A4+DCT is
free at both capability levels, with retention \emph{improving} as the policy strengthens
(98.4\%$\rightarrow$99.5\%; 95.0\% vs.\ 95.5\% bf16 on the strong checkpoint). W3A8 is
not: free on the weak checkpoint, it loses 5 points on the strong one---compression
headroom shrinks as policy quality rises, so recipes must be validated on strong
checkpoints. Among 16 candidate transforms, one principle governs all outcomes: global
dense mixers spread energy and work; PCA concentrates energy and is dangerous. The
execution-horizon study adds an orthogonal deployment rule: truncating action chunks
collapses success identically at every precision---chunked execution is load-bearing,
quantization is free, and full-chunk W4A4+DCT is simultaneously the accuracy and compute
optimum. Real Jetson AGX Orin deployment confirms 97.5\% retention at 86.0\% closed-loop
success---the first real-hardware closed-loop result for a quantized L1-regression VLA,
running at $\approx$24\,Hz within a 4.1\,GB memory footprint.

% ============================================================================

\begin{thebibliography}{99}
\small

\bibitem{openvla}
  A.~Kim et al., ``OpenVLA: An Open-Source Vision-Language-Action Model,''
  \textit{arXiv:2406.09246}, 2024.

\bibitem{openvlaoft}
  K.~Pertsch et al., ``Fast Robot Learning with Vision-Language Models (OpenVLA-OFT),''
  \textit{arXiv:2409.12191}, 2024.

\bibitem{libero}
  B.~Liu et al., ``LIBERO: Benchmarking Knowledge Transfer for Lifelong Robot Learning,''
  in \textit{NeurIPS}, 2023.

\bibitem{gptq}
  E.~Frantar et al., ``GPTQ: Accurate Post-Training Quantization for Generative Pre-trained
  Transformers,'' in \textit{ICLR}, 2023.

\bibitem{awq}
  J.~Lin et al., ``AWQ: Activation-aware Weight Quantization for LLM Compression and
  Acceleration,'' in \textit{MLSys}, 2024.

\bibitem{smoothquant}
  G.~Xiao et al., ``SmoothQuant: Accurate and Efficient Post-Training Quantization for
  Large Language Models,'' in \textit{ICML}, 2023.

\bibitem{quarot}
  S.~Ashkboos et al., ``QuaRot: Outlier-Free 4-Bit Inference in Rotated LLMs,''
  \textit{arXiv:2404.00456}, 2024.

\bibitem{spinquant}
  Z.~Liu et al., ``SpinQuant: LLM Quantization with Learned Rotations,''
  in \textit{ICLR}, 2025.

\bibitem{quip}
  J.~Chee et al., ``QuIP\#: Even Better LLM Quantization with Hadamard Incoherence and
  Lattice Codebooks,'' \textit{arXiv:2402.04396}, 2024.

\bibitem{sun2024massive}
  M.~Sun et al., ``Massive Activations in Large Language Models,''
  \textit{arXiv:2402.17762}, 2024.

\bibitem{pi0}
  K.~Black et al., ``$\pi_0$: A Vision-Language-Action Flow Model for General Robot
  Control,'' \textit{arXiv:2410.24164}, 2024.

\bibitem{groot}
  NVIDIA, ``GR00T N1: An Open Foundation Model for Generalist Humanoid Robots,''
  \textit{arXiv:2503.14734}, 2025.

\bibitem{qvla}
  Y.~Xu et al., ``QVLA: Not All Channels Are Equal in Vision-Language-Action Model's
  Quantization,'' \textit{arXiv:2602.03782}, 2026.

\bibitem{quantvla}
  J.~Zhang et al., ``QuantVLA: Scale-Calibrated Post-Training Quantization for
  Vision-Language-Action Models,'' \textit{arXiv:2602.20309}, 2026.

\bibitem{omegaqvla}
  X.~Wang et al., ``$\Omega$-QVLA: Robust Quantization for Vision-Language-Action Models
  via Composite Rotation and Per-step Scaling,'' \textit{arXiv:2605.28803}, 2026.

\bibitem{dyqvla}
  Z.~Zheng et al., ``DyQ-VLA: Temporal-Dynamic-Aware Quantization for Embodied
  Vision-Language-Action Models,'' \textit{arXiv:2603.07904}, 2026.

\end{thebibliography}

\end{document}
